A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-04-04 intro.tex
working...

    The number of analysis tools is expected to grow in the future as the scope of MDO problems continues
    to evolve, and as computing is increasingly commoditized. This expansion of scope will be driven, in
    part, by consideration of additional disciplines. Current trends on the horizon for aircraft MDO studies
    include incorporation of manufacturing analyses [[REF]], subsystem performance [[REF]], and fleet-level
    aggregate models of emissions, noise, or economics [[REF]].

@@

    Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
    data flows to assist an engineering team conducting this task. Because the problem deals with
    interconnectivity, we base our approach on the representations and techniques of graph theory. The 
    proposed graph syntax is closely related to the REMS system proposed by 
    Alexandrov and Lewis\cite{alexandrov2004}. The approach taken here should be viewed as an extention 
    of the concepts proposed in REMS with the goal of allowing the graphs to more closely interact with 
    existing design processes and MDAO frameworks. The approach begins by 
    constructing the \emph{maximal connectivity graph (MCG)} describing all possible
    interconnections between the analysis tools proposed by the engineers. Graph operations are then
    conducted reduce the MCG down to a \emph{fundamental problem graphs (FPG)} that describes the minimum set of analysis
    tools needed to solve the specified system-level design problem. The FPG does not predispose
    any particular solution procedure for the problem; any relevant MDO solution architecture, e.g. MDF,
    IDF, CO, BLISS, could be selected \emph{post facto} to implement the system-level optimization by
    constructing a \emph{problem solution graph (PSG)} and the corresponding problem solution approach.
    The concept of the FPG and the identification of feasible FPGs are the main contributions of the paper.

    The paper is organized as follows. First, we describe the differences between a fundamental problem
    formulation, which is based only on the system-level optimization problem statement that the
    engineers desire to solve and on the available analysis tools, and a specific problem formulation, which
    additionally presumes a specific solution to the problem. Next, we survey the literature in applying
    graph theoretic and formal language approaches to multidisciplinary design problem formulation. 
    We then discuss our graph syntax and representation of MDO problems and describe the procedures for 
    determining the MCG and FPG. Finally, we present an example problem based on an MDO analysis of a 
    commercial aircraft and discuss additional applications enabled by our approach.

\section{Specific vs Fundamental Problem Formulation }

@@

    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
    and $z$ will be iterated on to convergence with $z^t$. Now the problem can be solved by running $B$ first and then $A$.
    Since the formulations in Eqs. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem and neither can be the
    FPF. They are both specific versions of the same more fundamental description of 
    the problem that is common between them. We present the FPF as follows:

@@

    The FPF in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} and Eq. \ref{eqn:simple2} because it has 
    two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
    either $A$ or $B$ could be run first or both could be run simultaneously. Alexandrov and Lewis investigated problem 
    formulation in this fully decoupled way as part of their work on 
    Reconfigurability in MDO Problem Synthesis (REMS) \cite{alexandrov2004}.
    In this work they show how removing either constraint and replacing 
    it with a direct dependence between the two analyses, a process they term ``collapsing'', 
    you could regain the earlier two problem formulations.

@@

    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
    it can be represented in an equivalent directed graph where nodes represent analysis tools and 
    edges represent information dependence between those tools. The ordering of elements in a DSM can be used to indicate 
    execution order.  For more complex problems, choosing the proper order to run analysis tools is a non-trival task. 
    Rogers et. al developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
    reduces the cost of solving highly coupled systems\cite{Rogers1996}. This re-ordering is done through 
    row operations on the DSM matrix and yields multiple specific problem 
    formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
    the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself.

>>>>

    The number of analysis tools is expected to grow in the future as the scope of MDO problems continues
    to evolve, and as computing is increasingly commoditized. This expansion of scope will be driven, in
    part, by consideration of additional disciplines. Current trends on the horizon for aircraft MDO studies
    include incorporation of manufacturing analyses [[REF]], subsystem performance [[REF]], and fleet-level
    aggregate models of emissions, noise, or economics [[REF]]. A recent survey by Shan and Wang covered 
    a range of different techniques proposed to solve High Dimentional, Expensive, Black-box (HEB) problems. 
    The large amount of relevant research they identified demonstrates the growing interest large scale 
    MDO problems that arise from multiple disciplines\cite{Shan2010}.

@@

    Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
    data flows to assist an engineering team conducting this task. Because the problem deals with
    interconnectivity, we base our approach on the representations and techniques of graph theory. The 
    proposed graph syntax is similar to the REMS system proposed by 
    Alexandrov and Lewis\cite{alexandrov2004}. The approach taken here could be viewed as an extention 
    of the concepts proposed in REMS with the goals of providing greater insight into design problem 
    structure.  The approach begins by fconstructing the \emph{maximal connectivity graph (MCG)} describing all possible
    interconnections between the analysis tools proposed by the engineers. Graph operations are then
    conducted reduce the MCG down to a \emph{fundamental problem graphs (FPG)} that describes the minimum set of analysis
    tools needed to solve the specified system-level design problem. The FPG does not predispose
    any particular solution procedure for the problem; any relevant MDO solution architecture, e.g. MDF,
    IDF, CO, BLISS, could be selected \emph{post facto} to implement the system-level optimization by
    constructing a \emph{problem solution graph (PSG)} and the corresponding problem solution approach.
    The concept of the FPG and the identification of feasible FPGs are the main contributions of the paper.
    For completeness we include node and edge types for building a PSG, but the details of how a given 
    problem can be decomposed are beyond the scope of this work.

    The paper is organized as follows. First, we describe the differences between 
    a fundamental problem formulation based only on the design problem
    that the engineers desire to solve and on the available analysis tools, 
    and a specific problem formulation, which additionally 
    presumes a specific solution to the problem. Next, we survey the literature in applying
    graph theoretic and formal language approaches to multidisciplinary design 
    problem formulation. We then discuss our graph syntax and representation of 
    MDO problems and describe how different structures within a graph map to important
    features of an MDAO problem. Next we present a procedure for determining possible 
    FPG from a given MCG. Finally, we present the graph based formulation of commercial 
    aircraft design problem and discuss additional applications enabled by our approach.

\section{Specific vs Fundamental Problem Formulation }
    \label{sec:spg_vs_fpg}

@@

    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. 
    $A$ is now dependent on the output of $B$, and $z$ will be iterated on to 
    convergence with $z^t$. Now the problem can be solved by running $B$ first 
    and then $A$.Since the formulations in Eqs. \ref{eqn:simple} and 
    \ref{eqn:simple2} both describe the same problem then neither can be the
    FPF. They are both specific versions of a more fundamental description of 
    the problem that is common between them. We present the fundamental 
    formulation as follows:

@@

    The formulation in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} 
    and Eq. \ref{eqn:simple2} because it has two constraints which both must be 
    met. The presence of both of these constraints fully decouples the problem so that 
    either $A$ or $B$ could be run first or both could be run simultaneously. 
    Alexandrov and Lewis, in their work with REMS, assert that fully decoupling 
    the analyses is key to achieving reconfigurability in the problem 
    formulation\cite{alexandrov2004}.

@@

    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
    it can be represented in an equivalent directed graph where nodes represent analysis tools and 
    edges represent information dependence between those tools. The ordering of elements in a DSM can be used to indicate 
    execution order.  For more complex problems, choosing the proper order to run analysis tools is a non-trival task. 
    Rogers et. al developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
    reduces the cost of solving highly coupled systems\cite{Rogers1996,rogers1996demaid}. This re-ordering is done through 
    row operations on the DSM matrix and yields multiple specific problem 
    formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
    the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself.
